---
title: "CellChat comparison NAFLD vs. healthy controls"
author: "Hilmar Berger"

output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---

```{r, message=FALSE}
library(CellChat)
library(ggplot2)                  
library(patchwork)
library(igraph)
suppressPackageStartupMessages(library(ComplexHeatmap))

```

```{r}
cellchat.healthy <- readRDS("complete_Ramachandran_NAFLD_secreted_no_mobile_healthy_only_v2/CellChat_NET_final_complete_Ramachandran_NAFLD_secreted_no_mobile_healthy_only_v2.rds")
cellchat.healthy <- updateCellChat(cellchat.healthy)
#> Update slot 'var.features' from a vector to a list
cellchat.nafld <- readRDS("complete_Ramachandran_NAFLD_secreted_no_mobile_diseased_only_v2/CellChat_NET_final_complete_Ramachandran_NAFLD_secreted_no_mobile_diseased_only_v2.rds")
cellchat.nafld <- updateCellChat(cellchat.nafld)
#> Update slot 'var.features' from a vector to a list
```


```{r}
# Define the cell labels to lift up
group.new = levels(cellchat.healthy@idents)
cellchat.nafld <- liftCellChat(cellchat.nafld, group.new)  

#> The CellChat object will be lifted up using the cell labels FIB-A, FIB-B, FIB-P, DC, Pericyte, MYL, Immune, ENDO, Muscle, MELA, Basal-P, Basal, Spinious
#> Update slots object@net, object@netP, object@idents in a single dataset...
object.list <- list(Healthy = cellchat.healthy, NAFLD = cellchat.nafld)
cellchat <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
#> Warning in mergeCellChat(object.list, add.names = names(object.list),
#> cell.prefix = TRUE): Prefix cell names!
#> Merge the following slots: 'data.signaling','net', 'netP','meta', 'idents', 'var.features' , 'DB', and 'LR'.
```

# Overview

```{r, fig.height=3, fig.width=6}
gg1 <- compareInteractions(cellchat, show.legend = F, group = c(1,2))
gg2 <- compareInteractions(cellchat, show.legend = F, group = c(1,2), measure = "weight")
gg1 + gg2
```

# Comparison NAFLD vs. Healthy




```{r, fig.height=8, fig.width=10, eval=FALSE, echo=FALSE}
gg1 <- rankNet(cellchat, mode = "comparison", stacked = T, do.stat = TRUE)
gg2 <- rankNet(cellchat, mode = "comparison", stacked = F, do.stat = TRUE)
gg1 + gg2
```

## Different outgoing/incoming interaction strength per celltype

```{r, fig.height=8, fig.width=10}
num.link <- sapply(object.list, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(object.list)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list[[i]], title = names(object.list)[i], weight.MinMax = weight.MinMax)
}
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
patchwork::wrap_plots(plots = gg)
```

## Number and strength of incoming/outgoing interactions for selected populations

Shown are interactions of adipose tissue MoMF and liver hepatocytes.

```{r, fig.height=8, fig.width=10, eval=FALSE}
weight.max <- getMaxWeight(object.list, attribute = c("idents","count"))
par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Outgoing number of interactions - ", names(object.list)[i]), sources.use = "Perivascular MF_Adipose")
}

for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Incoming number of interactions - ", names(object.list)[i]), targets.use = "Perivascular MF_Adipose")
}

for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Outgoing number of interactions - ", names(object.list)[i]), sources.use = "KC-like_Adipose")
}

for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Incoming number of interactions - ", names(object.list)[i]), targets.use = "KC-like_Adipose")
}


for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Outgoing number of interactions - ", names(object.list)[i]), sources.use = "Lipid Associated MF_Adipose")
}

for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Incoming number of interactions - ", names(object.list)[i]), targets.use = "Lipid Associated MF_Adipose")
}

for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Outgoing number of interactions - ", names(object.list)[i]), sources.use = "Hepatocytes_Liver")
}

for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Incoming number of interactions - ", names(object.list)[i]), targets.use = "Hepatocytes_Liver")
}
```

### Comparison of interactions per pathway and cell type

#### Outgoing

```{r, fig.height=10, fig.width=10}
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i], width = 10, height = 14)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i+1], width = 10, height = 14)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```

#### Incoming 

```{r, fig.height=10, fig.width=10}
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "incoming", signaling = pathway.union, title = names(object.list)[i], width = 10, height = 14, color.heatmap = "GnBu")
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "incoming", signaling = pathway.union, title = names(object.list)[i+1], width = 10, height = 14, color.heatmap = "GnBu", )
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```


#### Incoming + Outgoing

```{r, fig.height=10, fig.width=10}
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathway.union, title = names(object.list)[i], width = 10, height = 14, color.heatmap = "OrRd")
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathway.union, title = names(object.list)[i+1], width = 10, height = 14, color.heatmap = "OrRd")
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```


# Differential interactions NAFLD vs. healthy (ligand-receptor pairs) by adipose cell type

## Perivascular MF

### Outoging

```{r, fig.width=14,fig.height=8}
netVisual_bubble(cellchat, sources.use = "Perivascular MF_Adipose", targets.use = c(5,6,9,10,11,12,14,17,18,19,21,22:27),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
```

### Incoming

```{r, fig.width=14,fig.height=8}
netVisual_bubble(cellchat, targets.use = "Perivascular MF_Adipose", sources.use = c(5,6,9,10,11,12,14,17,18,19,21,22:27),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
```

## DC-like

### Outgoing

```{r, fig.width=14,fig.height=8}
netVisual_bubble(cellchat, sources.use = "DC-like_Adipose", targets.use = c(5,6,9,10,11,12,14,17,18,19,21,22:27),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
```

### Incoming

```{r, fig.width=14,fig.height=8}
netVisual_bubble(cellchat, targets.use = "DC-like_Adipose", sources.use = c(5,6,9,10,11,12,14,17,18,19,21,22:27),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
```


## Lipid Associated MF

### Outgoing

```{r, fig.width=14,fig.height=8}
netVisual_bubble(cellchat, sources.use = "Lipid Associated MF_Adipose", targets.use = c(5,6,9,10,11,12,14,17,18,19,21,22:27),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
```

### Incoming

```{r, fig.width=14,fig.height=8}
netVisual_bubble(cellchat, targets.use = "Lipid Associated MF_Adipose", sources.use = c(5,6,9,10,11,12,14,17,18,19,21,22:27),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
```




# Individual pathways

## Hierarchical plots

Plots show selected cell populations (here: adipose tissue MF populations) on the left and all other populations on the right of each plot. Dots in the middle denote the targets (receivers) of the signals, while all other dots on the right and left are senders. While in the left figure the selected populations are receivers of signals, on the right the remaining populations are receivers. Self-self connections denote likely paracrine loops. 

```{r, fig.height=10, fig.width=18}
# Hierarchy plot
pathways.show <- c("VEGF","PDGF",  "VISFATIN","CXCL","CCL","CALCR","BAG","IGF","GALECTIN","IL16","PTN","MIF","ANGPTL","ANNEXIN","TNF", "TGFb") 
for (pp in pathways.show) {
  weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pp) # control the edge weights across different datasets
  vertex.receiver = c(4,7,8,16) # Left portion of hierarchy plot the shows signaling to dermal cells and right portion shows signaling to epidermal cells
  par(mfrow = c(1,2), xpd=TRUE)
  for (i in 1:length(object.list)) {
    netVisual_aggregate(object.list[[i]], signaling = pp, vertex.receiver = vertex.receiver, edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pp, names(object.list)[i]), layout="hierarchy", title.space = 1, vertex.weight.max = 2)
  }  
}

```



```{r, fig.height=8, fig.width=10, eval=FALSE}
# Circle plot
pathways.show <- c("VEGF") 
weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pathways.show) # control the edge weights across different datasets
par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list)) {
  netVisual_aggregate(object.list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]))
}
```


## Circos plots

```{r, fig.height=10, fig.width=20}
# Chord diagram
#pathways.show <- c("VEGF", "GAS","IL1", "CD40","TNF") 
for (pp in pathways.show) {
  par(mfrow = c(1,2), xpd=FALSE)
  for (i in 1:length(object.list)) {
    p = netVisual_aggregate(object.list[[i]], signaling = pp, layout = "chord", signaling.name = paste(pp, names(object.list)[i]))
  }
}

```

# Genes in differential selected pathways

Green = healthy adipose/liver  
Red = NAFLD adipose/cirrhotic liver


```{r, fig.height=8, fig.width=14, warning=FALSE, message=FALSE}
#cellchat@meta$datasets = factor(cellchat@meta$datasets, levels = c("Healthy","NAFLD")) # set factor level
for (pp in pathways.show) {
  p = plotGeneExpression(cellchat, signaling = pp, split.by = "datasets", colors.ggplot = T, color.use = c("green","red") )
  print(p)
}
```



<!-- # Top 20% of differential interactions -->

<!-- ```{r, fig.width=14, fig.height=8} -->
<!-- par(mfrow = c(1,2), xpd=TRUE) -->
<!-- netVisual_diffInteraction(cellchat, weight.scale = T, top=0.2) -->
<!-- netVisual_diffInteraction(cellchat, weight.scale = T, measure = "weight", top=0.2) -->
<!-- ``` -->

<!-- ```{r, fig.width=14, fig.height=8, eval=FALSE} -->
<!-- gg1 <- netVisual_heatmap(cellchat) -->
<!-- #> Do heatmap based on a merged object -->
<!-- gg2 <- netVisual_heatmap(cellchat, measure = "weight") -->
<!-- #> Do heatmap based on a merged object -->
<!-- gg1 + gg2 -->
<!-- ``` -->


<!-- ```{r, eval=FALSE} -->
<!-- weight.max <- getMaxWeight(object.list, attribute = c("idents","count")) -->
<!-- par(mfrow = c(1,2), xpd=TRUE) -->
<!-- for (i in 1:length(object.list)) { -->
<!--   netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(object.list)[i])) -->
<!-- } -->
<!-- ``` -->

# Software versions

```{r}
sessionInfo()
```

