---
title: "CellChat analysis of cluster in all NAFLD samples"
author: "Hilmar Berger"

output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---


```{r}
library(Seurat)
library(ggplot2)

library(CellChat)
library(patchwork)

library(future)
future::plan("multiprocess", workers = 4) # do parallel

#sample="Liver myeloid CCl4 3Weeks_JH190129/JH190129"
sample = "complete_NAFLD_MoMF.DC"
output_folder = make.names(sample)

if(!dir.exists(output_folder)) dir.create(output_folder)
```

```{r}
CellChatDB <- CellChatDB.human # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)
```

```{r}
dplyr::glimpse(CellChatDB$interaction)
```

# Overview 

This analysis aims at identifying cell-cell communication using the CellChat method. We follow the guidelines from https://github.com/sqjin/CellChat/blob/master/tutorial/CellChat-vignette.html.

We run the analysis for a single sample only (using clusters from the full normal diets + CCl4 set). The full ligand-receptor interaction database is used. 

## Preparation and inference

Cell-cell-interactions are inferred at the gene and pathway level. 

```{r}
set.seed(1234)
reload_data = T

stage1_file = file.path(output_folder, paste0("CellChat_stage1_",make.names(sample),".rds"))
if (reload_data) {
  load("../Analysis/Results/data_storage/2021-05-31/BCN_NAFLD_MoMF_DC_only.combined_cluster_annotated.RData")

  #scData_MoMF.DC$group_by_batch = paste0(scData_MoMF.DC$group,"_", scData_MoMF.DC$batch)
  #scData = subset(scData_MoMF.DC, group_by_batch == sample)
  scData = scData.combined
  
  cluster_counts = table(scData$cluster_label)
  good_clusters = names(cluster_counts[cluster_counts > 30])
  
  scData = subset(scData, cluster_label %in% good_clusters)
  scData = DietSeurat(scData)
  
  data.input = scData@assays$RNA@data
  meta = scData@meta.data[, c("cluster_label"),drop=F]
  
  cellchat <- createCellChat(object = data.input, meta=meta, group.by="cluster_label")
  
  #CellChatDB.use <- CellChatDB 
  #CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling")
  CellChatDB.use <- CellChatDB
  cellchat@DB <- CellChatDB.use

  # subset the expression data of signaling genes for saving computation cost
  cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database

  cellchat <- identifyOverExpressedGenes(cellchat)
  cellchat <- identifyOverExpressedInteractions(cellchat)
  
  # project gene expression data onto PPI network (optional)
  cellchat <- projectData(cellchat, PPI.human)
  
  cellchat <- computeCommunProb(cellchat)
  # Filter out the cell-cell communication if there are only few number of cells in certain cell groups
  cellchat <- filterCommunication(cellchat, min.cells = 10)

  cellchat <- computeCommunProbPathway(cellchat)

  cellchat <- aggregateNet(cellchat)

  saveRDS(cellchat, file = stage1_file)  
} else {
  cellchat = readRDS(stage1_file)
}
```

```{r}
df.net <- subsetCommunication(cellchat)
df.net.p <- subsetCommunication(cellchat, slot.name="netP")

write.table(df.net, file=file.path(output_folder, "Interactions_gene_level.txt"), sep="\t", quote=F, row.names=F)

write.table(df.net.p, file=file.path(output_folder, "Interactions_pathway_level.txt"), sep="\t", quote=F, row.names=F)
```


## Identified interactions - overview

```{r, fig.height=6, fig.width=10}
tmp = table(df.net.p$pathway)
barplot(sort(tmp), las=2, main="Count of identified pathway interactions across all clusters")
```


# Aggregated cell-cell communication network

We here look at the number of links between clusters and summarized communication probabilities for each cluster. 

## Globally

```{r, fig.width=10, fig.height=7}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

## For each cluster separately

```{r, fig.width=10, fig.height=10}
mat <- cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

## For each pathway separately

```{r, fig.width=8, fig.height=8}
pathways.show <- cellchat@netP$pathways

par(mfrow=c(1,1))
for (pp in pathways.show) {
  netVisual_aggregate(cellchat, signaling = pp, layout = "chord")  
}

```


```{r, eval=FALSE}
#pathways.show <- c("WNT") 
# Hierarchy plot
# Here we define `vertex.receive` so that the left portion of the hierarchy plot shows signaling to fibroblast and the right portion shows signaling to immune cells 
vertex.receiver = seq(1,4) # a numeric vector. 
netVisual_aggregate(cellchat, signaling = pathways.show,  vertex.receiver = vertex.receiver)
```


```{r, fig.width=10, fig.height=8}
# Heatmap
par(mfrow=c(1,1))
for (pp in pathways.show) {
  netVisual_heatmap(cellchat, signaling = pp, color.heatmap = "Reds")
  netAnalysis_contribution(cellchat, signaling = pp)
}

```


```{r, eval=FALSE}
# Access all the signaling pathways showing significant communications
pathways.show.all <- cellchat@netP$pathways
# check the order of cell identity to set suitable vertex.receiver
levels(cellchat@idents)
#vertex.receiver = seq(1,4)
for (i in 1:length(pathways.show.all)) {
  # Visualize communication network associated with both signaling pathway and individual L-R pairs
  #netVisual(cellchat, signaling = pathways.show.all[i], vertex.receiver = vertex.receiver, layout = "hierarchy")
  # Compute and visualize the contribution of each ligand-receptor pair to the overall signaling pathway
  gg <- netAnalysis_contribution(cellchat, signaling = pathways.show.all[i])
  ggsave(filename=file.path(output_folder, paste0(pathways.show.all[i], "_L-R_contribution.pdf")), plot=gg, width = 3, height = 2, units = 'in', dpi = 300)
}
```

```{r, fig.width=16, fig.height=15}
# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
netVisual_bubble(cellchat, remove.isolate = FALSE)
#> Comparing communications on a single object
```

```{r, fig.width=15, fig.height=15, eval=FALSE}
# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
# show all the interactions sending from Inflam.FIB
netVisual_chord_gene(cellchat, lab.cex = 0.5,legend.pos.y = 30)
```

<!-- # WNT pathway member expression -->

<!-- ```{r,fig.width=10, fig.height=10} -->
<!-- plotGeneExpression(cellchat, signaling = "WNT") -->
<!-- ``` -->

# Signaling roles 

```{r,fig.width=10, fig.height=4}
# Compute the network centrality scores
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # the slot 'netP' means the inferred intercellular communication network of signaling pathways
# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(cellchat, signaling = pathways.show, width = 8, height = 2.5, font.size = 10)
```

<!-- ## Dominant senders and receivers -->

<!-- ```{r,fig.width=10, fig.height=10} -->
<!-- # Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways -->
<!-- gg1 <- netAnalysis_signalingRole_scatter(cellchat) -->
<!-- #> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways -->
<!-- # Signaling role analysis on the cell-cell communication networks of interest -->
<!-- gg2 <- netAnalysis_signalingRole_scatter(cellchat, signaling = c("WNT","NOTCH","BMP")) -->
<!-- #> Signaling role analysis on the cell-cell communication network from user's input -->
<!-- gg1 + gg2 -->
<!-- ``` -->

```{r,fig.width=10, fig.height=8}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing")
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming")
ht1 + ht2
```



# Global communication patterns


```{r}
library(NMF)
library(ggalluvial)
```


## Outgoing patterns


```{r}
selectK(cellchat, pattern = "outgoing")
```

```{r, fig.height=6, fig.width=10}
nPatterns = 3 # ADJUSTED
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatterns)
```

```{r, fig.height=12, fig.width=12}
# river plot
netAnalysis_river(cellchat, pattern = "outgoing")
```

```{r, fig.height=12, fig.width=12}
netAnalysis_dot(cellchat, pattern = "outgoing")
```


## Incoming patterns


```{r}
selectK(cellchat, pattern = "incoming")
```

```{r, fig.height=6, fig.width=10}
nPatterns = 4 # ADJUSTED
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatterns)
```

```{r, fig.height=12, fig.width=12}
# river plot
netAnalysis_river(cellchat, pattern = "incoming")
```

```{r, fig.height=12, fig.width=12}
netAnalysis_dot(cellchat, pattern = "incoming")
```

```{r}
saveRDS(cellchat, file = file.path(output_folder, paste0("CellChat_NET_final_",make.names(sample),".rds") ) )
```

# Software versions

```{r}
sessionInfo()
```

