---
title: "Cluster identification"
author: "Hilmar Berger"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---


```{r, message=FALSE}
rm(list=ls())
library(Seurat)
library(ggplot2)

library(readxl)
library(writexl)
library(knitr)
library(data.table)
library(cowplot)

library(pheatmap)


result_folder = paste("./Results",format(Sys.time(), "%Y-%m-%d"),sep="/")
if (!file.exists(result_folder)) dir.create(result_folder, recursive=T)

data_storage_folder = file.path("./Results/data_storage","2021-05-31")
if (!file.exists(data_storage_folder)) dir.create(data_storage_folder, recursive=T)

library(future)
plan("multisession", workers = 2)

rerun_analyses = F
```

```{r}
load(file.path(data_storage_folder, "BCN_NAFLD.combined.RData"))
set.seed(1234)
#scData.combined <- FindClusters(scData.combined, resolution = 0.8)
#gc()
```

# Cluster overview

```{r, fig.width=14, fig.height=8}
DimPlot(scData.combined, reduction = "umap", label = TRUE)
```

```{r, fig.width=14, fig.height=8}
DimPlot(scData.combined, reduction = "umap", label = TRUE, split.by = "sample_label")
```

```{r}
#DEN_CCL4_markers <- FindAllMarkers(object = scData.combined, only.pos = TRUE, min.pct = 0.6, min.diff.pct = 0.3, thresh.use = 0.3, test.use = "roc")
if(rerun_analyses) {
  cluster_markers <- FindAllMarkers(object = scData.combined, only.pos = TRUE)
  save(cluster_markers, file=file.path(data_storage_folder, "All_markers.Rdata"))
  write_xlsx(cluster_markers,file.path(result_folder, "All_markers.xlsx"))
} else {
  load(file.path(data_storage_folder, "All_markers.Rdata"))
}
```

<!-- ```{r} -->
<!-- marker_matrix = dcast(as.data.table(DEN_CCL4_markers), gene ~ cluster, value.var="avg_logFC") -->
<!-- tmp1 = as.matrix(marker_matrix[,2:ncol(marker_matrix)]) -->
<!-- # tmp = cor(tmp1, method="spearman", use="pairwise") -->
<!-- # pheatmap(tmp, cluster_rows=F, cluster_cols=F) -->

<!-- tmp1[is.na(tmp1)] <- 0 -->
<!-- tmp = cor(tmp1, method="spearman", use="pairwise") -->
<!-- pheatmap(tmp, cluster_rows=F, cluster_cols=F, main="Pairwise marker gene correlations") -->
<!-- ``` -->

```{r, eval=FALSE}
all_cluster_numbers = as.character(sort(unique(scData.combined$seurat_clusters)))
tmp = table(scData.combined$seurat_clusters)
all_cluster_numbers_gt_3_cells = names(tmp[tmp>3])
```


```{r, eval=FALSE}
if(rerun_analyses) {
DefaultAssay(scData.combined) <- "RNA"

  cons_markers = list()
  for (n in all_cluster_numbers_gt_3_cells) {
    nn = as.integer(n)
    cons_markers[[as.character(n)]] <- tryCatch({FindConservedMarkers(scData.combined, ident.1 = nn, grouping.var = "group", verbose = FALSE)}, error=function(e) NULL)
  }
  save(cons_markers, file=file.path(data_storage_folder,"Conserved_cluster_markers.Rdata"))
  
  tmp1 = mapply(function(x,y) {x$cluster=rep(y, nrow(x));x$GeneSymbol = rownames(x); x$avg_logFC_global=apply(x[,grepl("_avg_logFC",colnames(x)),drop=F],1,mean,na.rm=T);x}, cons_markers, names(cons_markers))
  all_unique_colnames = unique(unlist(lapply(tmp1, colnames)))
  tmp2 = lapply(tmp1, function(x) {missing_cols = all_unique_colnames[!all_unique_colnames %in% colnames(x)]; for (m in missing_cols) {x[[m]]=NA}; x})
  cons_markers_all = do.call(rbind, tmp2)
  write_xlsx(cons_markers_all,file.path(result_folder, "All_cons_markers.xlsx"))
  
} else {
  load(file.path(data_storage_folder,"Conserved_cluster_markers.Rdata"))
}

```


```{r}

plot_genes <- function(dataset, genes) {
  
  all_genes =rownames(dataset@assays$RNA)
  
  for (g in sel_markers) {
    if (!g %in% all_genes) {
      print(paste0("Gene ",g," has no data (possibly filtered out)."))
      next
    } else {
        p1  =FeaturePlot(object = dataset, features = g, reduction="umap", label = T)
        p2 = VlnPlot(dataset, features = g, ncol = 1, pt.size = 0.01)
        print(plot_grid(p1,p2, ncol=2,align="h"))
    }
  }
}
```



## Positive/Negative selection markers 

These are markers used in the cell sorting to negatively select cells for scRNA analysis. Negative markers should mainly label cells that should not have passed the selection procedure.

Cells were sorted as CD45+Ly6G-Tcrb-CD4-CD19-NK1.1- cells.

CD45 is Ptprc.  
Ly6g is transiently expressed in monocytes in BM, while it is increasingly expressed in granulocytes and neutrophils with maturation.  
Pglyrp1 - circulating polymorphonuclear leukocytes (granulocytes + mast cells)  

We here add Cd3e/d/g as pan-T marker and Cd20 (Ms4a1) as mature B-cell marker.
Gata3, Cxcr6, Rora - T-cell markers

Ly6a(Sca-1) - positive in HSC/MPP
Procr,Fdg5, Hoxb5 - HSC

Epor, Slamf1(Cd150) - MegE - lineage

```{r, fig.width=14, fig.height=6}

sel_markers=toupper(c("Ptprc","Tcrb","Ly6g","Pglyrp1","Cd4","Cd3e","Cd3d","Cd3g", "Gata3","Cxcr6","Rora","Cd7","Cd8a","Cd19","Ms4a1","Ly6a", "Cd34","Procr","Fgd5","Hoxb5","Epor","Slamf1","EPCAM","VIM"))
DefaultAssay(scData.combined) <- "RNA"

plot_genes(scData.combined, sel_markers)

```

## Granulocyte subpopulations

Pglyrp1 - circulating polymorphonuclear leukocytes (granulocytes + mast cells)  
Csf3r - colony stimulating factor 3 receptor (granulocytes)  G-CSF-R. Higher in mature blood neutrophils (Shinjo et al. 1995 doi: 10.1111/j.1365-2141.1995.tb05390.x)

Mpo,Elane - neutrophil markers  
Prg2,Prg3 - eosinophils (Dahlin)  
Gzmb, Cma1 - Mast cell (Dahlin)  
Ms4a2, Cpa3 Mast cell, basophil (Dahlin)  
Prss34, Mcpt8 - basophil (Dahlin)  
Cd200r3 is a receptor on mastcells on basophils.  
Fcer1a - basophils, mastcells - high affinity IgE receptor  
Ggnt2, Ear2, Siglecf - eosinophils  
Il5ra - eosinophil progenitor  
Chil3 - immature neutrophils (Nio et al 2004  doi: 10.1007/s00418-004-0654-4)  


```{r, fig.width=14, fig.height=6}

sel_markers=toupper(c("Pglyrp1","Csf3r","Mpo","Elane","Chil3","Cd200r3","Gngt2","Ear2","Il5ra","Siglecf","Prg2","Prg3","Gzmb","Cma1","Ms4a2","Cpa3","Prss34","Mcpt8","Fcer1a"))
DefaultAssay(scData.combined) <- "RNA"

plot_genes(scData.combined, sel_markers)

```


## MoMF markers

Adgre4  Blood monocytes (Caminschi et al 2001, doi: 10.4049/jimmunol.167.7.3570)
Adgre1 = F4/80 = Emr1, Blood Monocytes

CD14lowCD16high - non-classical MoMF    
CD16 - Fcgr4 (FCGR3A/B)  
Fcgr1 (Cd64) MoMF  
Spn (Cd45) classical Mo  
Cd11b = Itgam  

EMP (Mass et al.) Csf1r + Kit + CD45 low AA4.1(Cd93)+   

```{r, fig.width=14, fig.height=6}
sel_markers=toupper(c("Adgre4", "Adgre1","Lyz1","Lyz2","Ly6c1","Ly6c2","Ccr2","Ccr5","Arg1","Mgst1","Fn1","Cd74","Cd68","S100a8","S100a9","Cd14","Fcgr4","Cx3cr1","Timd4","Fcgr1","Spn","Itgam","Csf1r","Kit","Cd93","Il7r","Flt3","Cxcr4","TREM2"))
DefaultAssay(scData.combined) <- "RNA"

plot_genes(scData.combined, sel_markers)
```

## Kupffer cell markers

```{r, fig.width=14, fig.height=6}
sel_markers=toupper(c("C1qa","C1qb","Cd163","Vsig4","Mafb","Marco","Mrc1","Cd14","Gpr183", "Clec4f","Nr1h3","Id3","Id1","Spic","Timd4"))
DefaultAssay(scData.combined) <- "RNA"

plot_genes(scData.combined, sel_markers)
```



## DC markers

Xcr1 cDC  
Siglech pDC  
Irf8 pre-DC1  
Ccr7 cDC2  
Pclaf (2810417H13Rik) CMoP / pre-DC1  
Cd209a cDC2, pDC  
Tlr2 cDC  
Tlr4 cDC  

Socs2 Tissue migratory DC (Nirschl et al. 2017, 10.1016/j.cell.2017.06.016)  


```{r, fig.width=14, fig.height=6}
sel_markers=toupper(c( "Socs2", "Xcr1", "Siglech", "Cd209a", "Irf8", "Pclaf", "Tlr2","Tlr4","Tlr7","Tlr9","CD1C","CLEC4C"))
DefaultAssay(scData.combined) <- "RNA"

plot_genes(scData.combined, sel_markers)
```


## MHC-II

```{r, fig.width=14, fig.height=6}
#sel_markers=c("H2-Oa","H2-DMa","H2-DMb1","H2-DMb2","H2-Ob","H2-Ab1","H2-Aa","H2-Eb1","H2-Eb2")
sel_markers = c("HLA-A","HLA-B","HLA-C","HLA-D","HLA-E","HLA-F","HLA-G","HLA-DRA","HLA-DRB","HLA-DQA1","HLA-DQB1","HLA-DPA1","HLA_DPB1","HLA_DMA","HLA_DMB","HLA-DOA","HLA_DOB","MICA","MICB")
DefaultAssay(scData.combined) <- "RNA"

plot_genes(scData.combined, sel_markers)
```

## Adipocyte markers

Sun et al. Nature 2020 (https://doi.org/10.1038/s41586-020-2856-x)  
CIDEA, LEP, TCF21, ADIPOQ - Adipocytes  
Ly6a (LY6L), Cd34 - Preadipocytes   


Liu et al. Sci Data 2019 (https://doi.org/10.1038/sdata.2019.31)  
CD105(ENG), CD73(NT5E), CD90(THY1), CD59, CD44, CD29(ITGB1) - ADSC  

```{r, fig.width=14, fig.height=6}
sel_markers=toupper(c("CA4","CIDEA","TBX1","LEP","FABP4","TCF21","ADIPOQ","PLIN1", "LY6L","CD34", "ENG","NT5E","THY1","CD59","CD44","ITGB1"))
DefaultAssay(scData.combined) <- "RNA"

plot_genes(scData.combined, sel_markers)
```


## Proliferation

```{r, fig.width=14, fig.height=6}
sel_markers=toupper(c("Top2a","Mki67","Pcna"))
DefaultAssay(scData.combined) <- "RNA"

plot_genes(scData.combined, sel_markers)
```


<!-- # Apoptosis score and mitochondrial RNA -->

<!-- We use the HALLMARK_APOPTOSIS gene set from MSigDB to compute an apoptosis score as the mean expression across all signature genes per cell.  -->

<!-- ```{r} -->
<!-- library(fgsea) -->

<!-- gset_collections_human = list() -->

<!-- gsea_gmt_folder = "/home/hilmar/Work/References/genesets/MSigDB/v7.1/" -->
<!-- gsea_gmt_files = c("h.all.v7.1.symbols.gmt") # ,"c2.all.v7.1.symbols.gmt","c3.all.v7.1.symbols.gmt","c5.bp.v7.1.symbols.gmt","c6.all.v7.1.symbols.gmt","c7.all.v7.1.symbols.gmt" -->
<!-- names(gsea_gmt_files) = c("HALLMARK") -->
<!-- gset_collections_human = list() -->
<!-- for (n in names(gsea_gmt_files)) { -->
<!--   gset_collections_human[[n]] = gmtPathways(paste(gsea_gmt_folder, gsea_gmt_files[[n]], sep="/") ) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- all_genes =rownames(scData.combined@assays$RNA) -->
<!-- homologene_human_mouse = fread("/home/hilmar/Work/References/HomologousGenes/HomoloGene/build68/homologene_9606_10090.txt") -->
<!-- all_symbols = data.frame(GeneSymbol=unique(all_genes)) -->
<!-- a1 = merge(all_symbols, homologene_human_mouse[,c("entrez_1","symbol_1","symbol_2"), with=F], by.x="GeneSymbol", by.y="symbol_2", all.x=T, sort=F) -->

<!-- mouse_to_human=a1 -->
<!-- mouse_to_human = subset(mouse_to_human, !is.na(GeneSymbol) & !is.na(symbol_1) & !symbol_1 == "") -->
<!-- rownames(mouse_to_human) = mouse_to_human$GeneSymbol -->
<!-- ``` -->


<!-- ```{r} -->
<!-- sel_genesets = c("HALLMARK_APOPTOSIS","HALLMARK_E2F_TARGETS", "HALLMARK_MITOTIC_SPINDLE","HALLMARK_REACTIVE_OXYGEN_SPECIES_PATHWAY") -->
<!-- for (gg in sel_genesets) { -->
<!--   sig_genes_hs = gset_collections_human[[1]][[gg]] -->
<!--   sig_genes_mm = subset(mouse_to_human, symbol_1 %in% sig_genes_hs )$GeneSymbol -->
<!--   sel_mat = Assays(subset(scData.combined, features=sig_genes_mm), "RNA")@data -->
<!--   scData.combined = AddMetaData(scData.combined, col.name = gg, apply(sel_mat, 2, mean, na.rm=T) ) -->
<!--   print(head(sig_genes_mm)) -->
<!--   print(length(sig_genes_mm)) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r, fig.height=8 } -->
<!-- for (gg in sel_genesets) { -->
<!--   print(FeaturePlot(object = scData.combined,  features = gg, reduction="umap", label=T)) -->
<!--   print(VlnPlot(object = scData.combined, features=gg, ncol = 1, pt.size = 0.01)) -->
<!--   tmp = data.frame(score = unlist(scData.combined[[gg]]), batch= scData.combined$batch, group_label = scData.combined$group ) -->
<!--   print(ggplot(tmp, aes(fill=group_label, x=group_label, y = score)) + geom_boxplot() + theme(axis.text.x = element_text(angle=90, hjust=1)) + ggtitle(gg)) -->

<!-- } -->


<!-- ``` -->

<!-- ## S and G2M specific genes -->

<!-- ```{r} -->
<!-- # those are the mouse homologues of human cell cycle genes that come with Seurat  -->
<!-- cell_cycle_genes = read.table("../../CVC_SC/Analysis/regev_lab_cell_cycle_genes_OK_working_with_phase.txt", sep="\t", header=F, stringsAsFactors = F) -->
<!-- colnames(cell_cycle_genes) = c("Symbol","Phase") -->
<!-- s.genes = subset(cell_cycle_genes, Phase=="S")$Symbol -->
<!-- g2m.genes = subset(cell_cycle_genes, Phase=="G2M")$Symbol -->
<!-- gsets = list("S_genes" = s.genes, "G2M_genes" = g2m.genes) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for (gg in names(gsets)) { -->
<!--   sig_genes_mm = gsets[[gg]] -->
<!--   sel_mat = Assays(subset(scData.combined, features=sig_genes_mm), "RNA")@data -->
<!--   scData.combined = AddMetaData(scData.combined, col.name = gg, apply(sel_mat, 2, mean, na.rm=T) ) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r, fig.height=8 } -->
<!-- for (gg in names(gsets)) { -->
<!--   print(FeaturePlot(object = scData.combined,  features = gg, reduction="umap", label=T)) -->
<!--   print(VlnPlot(object = scData.combined, features=gg, ncol = 1, pt.size = 0.01)) -->
<!--   tmp = data.frame(score = unlist(scData.combined[[gg]]), batch= scData.combined$batch, group_label = scData.combined$group ) -->
<!--   print(ggplot(tmp, aes(fill=group_label, x=group_label, y = score)) + geom_boxplot() + theme(axis.text.x = element_text(angle=90, hjust=1)) + ggtitle(gg)) -->

<!-- } -->


<!-- ``` -->



## Mito score

```{r}
FeaturePlot(object = scData.combined,  features = "percent_mito", reduction="umap", label=T)
```

```{r, fig.height=8}
tmp = data.frame(percent_mito = scData.combined$percent_mito, batch= scData.combined$batch, group_label = scData.combined$group )
ggplot(tmp, aes(fill=batch, x=group_label, y = percent_mito)) + geom_boxplot() + theme(axis.text.x = element_text(angle=90, hjust=1))

```


## Scores and RNA count per cluster

```{r, fig.width=10, fig.height=6}
VlnPlot(scData.combined, features = "percent_mito", ncol = 1, pt.size = 0.01)
#VlnPlot(scData.combined, features = "HALLMARK_APOPTOSIS", ncol = 1, pt.size = 0.01)
VlnPlot(scData.combined, features = "nCount_RNA", ncol = 1, pt.size = 0.01)
VlnPlot(scData.combined, features = "nFeature_RNA", ncol = 1, pt.size = 0.01)
```

```{r}
FeaturePlot(object = scData.combined,  features = "nFeature_RNA", reduction="umap", label=T)
```


# Final cluster assigment

```{r}
cluster_assignment = read.table("Cluster_assignments/BCN_NAFLD_Clusters_2021-06-18.txt", sep="\t",header=T, stringsAsFactors = F)
options(width = 180)
print(cluster_assignment)
rownames(cluster_assignment) = as.character(cluster_assignment$ClusterID)
scData.combined = AddMetaData(scData.combined, cluster_assignment[as.character(scData.combined$seurat_clusters),"Label"], col.name = "cluster_label")
scData.combined = AddMetaData(scData.combined, cluster_assignment[as.character(scData.combined$seurat_clusters),"Type"], col.name = "cluster_type")
```

```{r, fig.width=14, fig.height=8}
scData.combined = SetIdent(scData.combined, value="cluster_label" )
p = DimPlot(scData.combined, reduction = "umap", label = TRUE)
print(p)
```

```{r}
pdf(file.path(result_folder, "UMAP_annotated.pdf"), width=14, height=8)
print(p)
dev.off()
```


```{r}
m = table(scData.combined$cluster_label, scData.combined$sample_label)
m
```

```{r, fig.width=10,fig.height=7}
mm  = as.matrix.data.frame(m)
colnames(mm) = colnames(m)
rownames(mm) = rownames(m)

mm_r = sweep(mm, 2, apply(mm,2,sum), "/")

pheatmap(mm_r, cluster_cols = F, cluster_rows=F)
```

## Known markers

CD14lowCD16high - non-classical MoMF    
CD16 - Fcgr4 (FCGR3A/B)  
Fcgr1 (Cd64) MoMF  
Spn (Cd43) classical Mo  
Cd11b = Itgam  

```{r, fig.height=10, fig.width=10}
sel_genes = toupper(c("Cd68","Csf1r","Cx3cr1","Cd14","Fcgr4","Itgam","Ccr2","Adgre4","Adgre1","Fcgr1","Spn","Ly6c","Ly6a"))

sel_mat = Assays(subset(scData.combined, features=sel_genes), "RNA")@data

sel_mat_ts = reshape2::melt(as.matrix(sel_mat))
colnames(sel_mat_ts) = c("Gene","CellID","Expression")
meta = scData.combined@meta.data
sel_mat_ts = merge(sel_mat_ts, meta[,c("cluster_label","group")], by.x="CellID",by.y=0, all.x=T, sort=F)

sel_mat_ts$Gene = gsub("Fcgr4","CD16",gsub("Itgam","CD11B",gsub("Adgre4","FIRE",gsub("Adgre1","F4/80",gsub("Fcgr1","CD64",gsub("Spn","CD43",sel_mat_ts$Gene))))))

ggplot(sel_mat_ts, aes(x=cluster_label,y=Expression, fill="black")) + geom_violin(scale = "width") + facet_grid(Gene ~ .) + theme(axis.text.x = element_text(angle=90, hjust=1)) + xlab("") 

```


```{r, fig.height=10, fig.width=10}
sel_genes = toupper(c("Cd68","CD3E","CD8","CD34","KRT8","KRT18","VIM", "ADIPOQ","NT5E","THY1","VWF"))

sel_mat = Assays(subset(scData.combined, features=sel_genes), "RNA")@data

sel_mat_ts = reshape2::melt(as.matrix(sel_mat))
colnames(sel_mat_ts) = c("Gene","CellID","Expression")
meta = scData.combined@meta.data
sel_mat_ts = merge(sel_mat_ts, meta[,c("cluster_label","group", "cluster_type")], by.x="CellID",by.y=0, all.x=T, sort=F)

#sel_mat_ts$Gene = gsub("Fcgr4","CD16",gsub("Itgam","CD11B",gsub("Adgre4","FIRE",gsub("Adgre1","F4/80",gsub("Fcgr1","CD64",gsub("Spn","CD43",sel_mat_ts$Gene))))))

ggplot(sel_mat_ts, aes(x=cluster_label,y=Expression, fill=cluster_type)) + geom_violin(scale = "width") + facet_grid(Gene ~ .) + theme(axis.text.x = element_text(angle=90, hjust=1)) + xlab("") 

```


```{r}
save(scData.combined,file=file.path(data_storage_folder, "BCN_NAFLD.combined_cluster_annotated.RData"))
```


# Export 

Cluster assignations have been exported to file *Metadata_BCN_Adipose_complete.txt*.

```{r}
tmp = scData.combined@meta.data
tmp$cell_id = rownames(tmp)
write.table( tmp , file="Metadata_BCN_Adipose_complete.txt", sep="\t", row.names = F, quote = F)
```



# Software versions

```{r}
sessionInfo()
```














