---
title: "Cluster markers - Myeloid cells only, RNA assay for DGE"
author: "Hilmar Berger"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---


```{r, message=FALSE}
rm(list=ls())
library(Seurat)
library(ggplot2)

library(readxl)
library(writexl)
library(knitr)
library(data.table)
library(cowplot)
library(magrittr)
library(dplyr)

library(pheatmap)

library(grid)
library(png)

library(clusterProfiler)

options(connectionObserver = NULL)
library(org.Hs.eg.db)
library(ReactomePA)
library(msigdbr)

result_folder = paste("./Results",format(Sys.time(), "%Y-%m-%d"),sep="/")
if (!file.exists(result_folder)) dir.create(result_folder, recursive=T)

data_storage_folder = file.path("./Results/data_storage","2021-05-31")
if (!file.exists(data_storage_folder)) dir.create(data_storage_folder, recursive=T)

library(future)
options("future.globals.maxSize"=2**33)
plan("multicore", workers = 8)


source("DotPlot_fixed.R")
source("DGE_Helper.R")

ref_folder = "../../References/"

rerun_analyses = F
preprocess_data = F
rerun_gsea_analyses = F
```

We only keep clusters annotated as MoMF, DC, and KC in the global analysis and reprocess the remaining cells using the same integrated analysis. 

```{r}
load(file.path(data_storage_folder, "BCN_NAFLD_MoMF_DC_only.combined_cluster_annotated.RData"))
```


# Cluster overview

```{r, fig.width=10, fig.height=6}
DimPlot(scData.combined, reduction = "umap", label = TRUE)
```

```{r, fig.width=14, fig.height=8}
DimPlot(scData.combined, reduction = "umap", label = TRUE, split.by = "sample_label")
```

# Marker genes for clusters

```{r}
#DEN_CCL4_markers <- FindAllMarkers(object = scData.combined, only.pos = TRUE, min.pct = 0.6, min.diff.pct = 0.3, thresh.use = 0.3, test.use = "roc")
if(rerun_analyses) {
  cluster_markers <- FindAllMarkers(object = scData.combined, only.pos = TRUE, assay = "RNA")
  save(cluster_markers, file=file.path(data_storage_folder, "All_markers_MoMF_DC_reprocessed_v2.Rdata"))
  write_xlsx(cluster_markers,file.path(result_folder, "All_markers_MoMF_DC_reprocessed_v2.xlsx"))
} else {
  load(file.path(data_storage_folder, "All_markers_MoMF_DC_reprocessed_v2.Rdata"))
}
```

```{r}
plot_top_genes <- function(data, dge_results_all, title) {
  for (n in sort(unique(data$cluster_label))) {
    tmp = subset(dge_results_all, cluster==n & p_val_adj < 0.05)
    tmp = tmp[order(-tmp$avg_log2FC),]
    if(nrow(tmp)<1) next
    tmp_up = subset(tmp, avg_log2FC > 0)
    if(nrow(tmp_up)<1) next
    tmp_up$direction = "up"
    #tmp_down = subset(tmp, avg_log2FC < 0)
    #tmp_down$direction = "down"
    
    #final_top_genes = rbind(head(tmp_up,10), head(tmp_down,10))
    final_top_genes = head(tmp_up, 10)
    final_top_genes = final_top_genes[order(final_top_genes$avg_log2FC),]
    final_top_genes$gene = factor(final_top_genes$gene, levels = final_top_genes$gene)
    p = ggplot(final_top_genes, aes(y=avg_log2FC, x=gene, fill=direction)) + geom_bar(stat="identity") + ggtitle(paste0(n, ", ", title)) + coord_flip() + theme(axis.text.y =  element_text(size=12) )
    print(p)
    
  }
}
```

```{r}
DefaultAssay(scData.combined) <- "RNA"
scData.combined <- ScaleData(scData.combined, verbose = TRUE, features = rownames(scData.combined@assays$RNA@counts))
gc()
```


## overview (10 top genes for each cluster)



```{r, fig.width=14, fig.height=12}
cluster_markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(scData.combined, features = top10$gene) + NoLegend()
```


## Percent expression of top markers across clusters

```{r, fig.width=16, fig.height=14}
top_5_markers = unique(unlist(tapply(cluster_markers$gene, cluster_markers$cluster, head, 5)))
DotPlot(scData.combined, features = top_5_markers , dot.scale = 4, cols = c("red", "blue"), assay = "RNA") +    RotatedAxis()
```

## Log2 Fold Changes of top 10 DE genes from each cluster

```{r, fig.width=12, fig.height=6}
plot_top_genes(scData.combined, cluster_markers, "Cluster markers")
```


<!-- ```{r} -->
<!-- marker_matrix = dcast(as.data.table(DEN_CCL4_markers), gene ~ cluster, value.var="avg_logFC") -->
<!-- tmp1 = as.matrix(marker_matrix[,2:ncol(marker_matrix)]) -->
<!-- # tmp = cor(tmp1, method="spearman", use="pairwise") -->
<!-- # pheatmap(tmp, cluster_rows=F, cluster_cols=F) -->

<!-- tmp1[is.na(tmp1)] <- 0 -->
<!-- tmp = cor(tmp1, method="spearman", use="pairwise") -->
<!-- pheatmap(tmp, cluster_rows=F, cluster_cols=F, main="Pairwise marker gene correlations") -->
<!-- ``` -->

```{r, eval=FALSE}
all_cluster_numbers = as.character(sort(unique(scData.combined$seurat_clusters)))
tmp = table(scData.combined$seurat_clusters)
all_cluster_numbers_gt_3_cells = names(tmp[tmp>3])
```


```{r, eval=FALSE}
if(rerun_analyses) {
DefaultAssay(scData.combined) <- "RNA"

  cons_markers = list()
  for (n in all_cluster_numbers_gt_3_cells) {
    nn = as.integer(n)
    cons_markers[[as.character(n)]] <- tryCatch({FindConservedMarkers(scData.combined, ident.1 = nn, grouping.var = "group", verbose = FALSE)}, error=function(e) NULL)
  }
  save(cons_markers, file=file.path(data_storage_folder,"Conserved_cluster_markers.Rdata"))
  
  tmp1 = mapply(function(x,y) {x$cluster=rep(y, nrow(x));x$GeneSymbol = rownames(x); x$avg_logFC_global=apply(x[,grepl("_avg_logFC",colnames(x)),drop=F],1,mean,na.rm=T);x}, cons_markers, names(cons_markers))
  all_unique_colnames = unique(unlist(lapply(tmp1, colnames)))
  tmp2 = lapply(tmp1, function(x) {missing_cols = all_unique_colnames[!all_unique_colnames %in% colnames(x)]; for (m in missing_cols) {x[[m]]=NA}; x})
  cons_markers_all = do.call(rbind, tmp2)
  write_xlsx(cons_markers_all,file.path(result_folder, "All_cons_markers.xlsx"))
  
} else {
  load(file.path(data_storage_folder,"Conserved_cluster_markers.Rdata"))
}

```


## Functional enrichment

We here use serveral data bases of known functionally relevant gene expression signatures and gene sets to identify those who are enriched among cluster marker lists. 

Data bases used are:

  - Molecular Signature DB (MSigDB, https://www.gsea-msigdb.org/gsea/msigdb), collections:
     - H (Hallmark), 
     - C2 (curated genesets and canonical pathways from KEGG, REACTOME etc.)
     - C3 (target genes of transcription factors and micro-RNAs)
     - C7 (immunologic signature gene sets)
  - KEGG DB (overlaps C2 database from MSigDB)
  - REACTOME DB (overlaps C2 database from MSigDB)
  - GO Biological Process
  <!-- - MAF_RES (Macrophage Resource)  - Genes up-/downregulated upon stimulation with different activators in macrophages (Xue et al, Immunity 2014, doi: 10.1016/j.immuni.2014.01.006) -->
  - Marker genes from clusters described in Hildgreth et al., Nat Immunology 2021, p.639


```{r}
signatures_gs <- import_sigs()
```


```{r}
title = "MoMF_DC_cluster_markers_v2"
project_name = "BCM_MarColl"
result_file = paste0("DGE_",title,"_",project_name,"_by_celltype_enrichment.Rdata")


if(rerun_gsea_analyses) {

  tmp = cluster_markers
  tmp$celltype = tmp$cluster
  tmp$GeneSymbol = tmp$gene
  all_enrichments = compute_functional_enrichment(scData.combined, tmp, title)
  
  save(all_enrichments, file=file.path(data_storage_folder, result_file))
} else {
  load(file.path(data_storage_folder, result_file))
}
```

### Top enriched terms per cluster

```{r, fig.width=18, fig.height=16}
for (n in sort(names(all_enrichments))) {
  p = my_dotplot(all_enrichments[[n]]) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(n)
  print(p)
}
```


<!-- ```{r, fig.height=10, fig.width=16} -->
<!-- for (n in names(all_enrichments)) { -->
<!--   p = dotplot(all_enrichments[[n]]) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(n) -->
<!--   print(p) -->
<!-- } -->
<!-- ``` -->


## Top 30 marker genes

Feature and violin plots for the top 30 marker genes of each cluster are written to file *Top30_Cluster_Markers.pdf*

```{r}

plot_genes2 <- function(dataset, genes) {
  
  all_genes =rownames(dataset@assays$RNA)
  
  for (g in sort(names(genes))) {
    if (!g %in% all_genes) {
      print(paste0("Gene ",g," has no data (possibly filtered out)."))
      next
    } else {
        p1  =FeaturePlot(object = dataset, features = g, reduction="umap", label = T, pt.size = 1, order=T) + ggtitle(paste0(g," (",genes[g],")"))
        p2 = VlnPlot(dataset, features = g, ncol = 1, pt.size = 0.01) + NoLegend() + ggtitle(paste0(g," (",genes[g],")"))
        print(plot_grid(p1,p2, ncol=2,align="h"))
    }
  }
}
```


```{r, fig.width=14, fig.height=6}
cluster_markers %>%
    group_by(cluster) %>%
    arrange(desc(avg_log2FC)) %>%
    top_n(n = 30, wt = avg_log2FC) -> top30

top30$Label = top30$cluster

marker_by_cluster = tapply(top30$Label, top30$gene, function(x) paste(unique(sort(x)), collapse=", "))

DefaultAssay(scData.combined) <- "RNA"
scData.combined$cluster_label = factor(scData.combined$cluster_label, levels=sort(unique(scData.combined$cluster_label)))
scData.combined <- SetIdent(scData.combined, value="cluster_label" )


png(file.path(result_folder, "Top30_Cluster_Markers_%04d.png"), width=1400, height=600)
plot_genes2(scData.combined, marker_by_cluster)
dev.off()

pngFiles = list.files(result_folder, pattern="*.png", full.names = T)
# code from https://jonkimanalyze.wordpress.com/2014/07/24/r-compile-png-files-into-pdf/
pdf(file.path(result_folder, "Top30_Cluster_Markers.pdf"), width = 14, height = 6)
n <- length(pngFiles)
for( i in 1:n) {
  pngFile <- pngFiles[i]
  pngRaster <- readPNG(pngFile)
  grid.raster(pngRaster)
  if (i < n) grid.newpage() #plot.new()
}
dev.off()

```



# Software versions

```{r}
sessionInfo()
```














