---
title: "BCN MarColl NAFLD scData - Preprocessing"
author: "Hilmar Berger"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---

```{r, message=FALSE}
library(Seurat)
library(ggplot2)

library(readxl)
library(knitr)
library(data.table)
library(cowplot)

library(pheatmap)

#library(future)
#plan("multiprocess", workers = 2)

data_storage_folder = file.path("./Results/data_storage","2021-05-31")
if (!file.exists(data_storage_folder)) dir.create(data_storage_folder, recursive=T)

```

```{r}
load(file.path(data_storage_folder, "All_Filtered_scData.Rdata"))

table(all_ge_data_filtered$group)
```
# Estimation and correction of cell cycle effect

We use a list of known cell cycle genes (provided by the Seurat package) to estimate the contribution of cell cycle phase on the global gene expression.

```{r}
# # those are the mouse homologues of human cell cycle genes that come with Seurat 
# cell_cycle_genes = read.table("../regev_lab_cell_cycle_genes_OK_working_with_phase.txt", sep="\t", header=F, stringsAsFactors = F)
# colnames(cell_cycle_genes) = c("Symbol","Phase")
#s.genes = subset(cell_cycle_genes, Phase=="S")$Symbol
#g2m.genes = subset(cell_cycle_genes, Phase=="G2M")$Symbol
s.genes = cc.genes.updated.2019$s.genes
g2m.genes = cc.genes.updated.2019$g2m.genes

```

```{r}
scData = all_ge_data_filtered
scData <- NormalizeData(object = scData, normalization.method = "LogNormalize", scale.factor = 1e4)
scData <- FindVariableFeatures(object = scData, selection.method = "vst")
scData <- ScaleData(object = scData, features = rownames(scData))
gc()
```

## Association of major variance components with cell cycle

```{r}
scData <- RunPCA(scData, features = VariableFeatures(scData), ndims.print = c(1:20), nfeatures.print = 20)
#scData <- RunPCA(scData, features = VariableFeatures(scData), ndims.print = c(1:100), nfeatures.print = 20)

# check top 100 positively loaded genes for PC_3 for cell cycle association in MSigDB
#xx = scData@reductions$pca@feature.loadings
#cat(paste(names(tail(sort(xx[, "PC_3"]), 100)), collapse="\n"))
gc()
```

<!-- There are some cell cycle markers in PCs 1 and 7 (Top2a, Mki67) and G2M markers in PC 13. -->

<!-- ```{r, fig.width=12} -->
<!-- DimHeatmap(scData, dims = c(1, 7, 13)) -->
<!-- ``` -->

## Cell cycle scoring

We use the Seura cell cycle scoring method to score association with S and G2/M phases and assign cells to G1, S or G2/M

```{r}
scData <- CellCycleScoring(scData, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

# view cell cycle scores and phase assignments
#head(scData[[]])
boxplot(scData$S.Score ~ scData$Phase)
boxplot(scData$G2M.Score ~ scData$Phase)
```

For most of the cells the S and G2/M scores are rather low, even for those assigned to the corresponding class. 

```{r, fig.height=6}
RidgePlot(scData, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), ncol = 4)
```

```{r, fig.width=13, fig.height=8}
# PCA on cell cycle genes reveals
scData <- RunPCA(scData, features = c(s.genes, g2m.genes))
DimPlot(scData) + geom_hline(yintercept = 0, lty=2) + geom_vline(xintercept = 0, lty=2)
DimPlot(scData,dims = c(2,3) )
DimHeatmap(scData, dims = c(1, 2))
```


```{r}
m = table(scData$Phase, scData$group)
kable(m)

tmp = as.matrix.data.frame(m)
colnames(tmp) = colnames(m)
rownames(tmp) = rownames(m)
mr = sweep(tmp, 2, apply(tmp,2,sum),"/")
kable(round(mr,2))

pheatmap(mr, scale="none", main="Relative contributions of cell cycle phases per experimental group", breaks=seq(0,1,by=0.01))
```
As can be expected, there is a larger proportion of G2M and S phase cells in the bone marrow samples. 

```{r}
#scData <- ScaleData(scData, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(scData))

# scData <- RunPCA(object = scData, pc.genes = scData@var.genes, do.print = F, pcs.compute = 100)
# #pdf(sprintf("%s/output/elbowplot.pdf",scData.workdir))
# ElbowPlot(object=scData,ndims=50)
# #dev.off()
# scData <- ProjectPCA(object = scData, do.print = F)
# scData <- FindClusters(object = scData, reduction.type = "pca", dims.use = 1:75, resolution = 0.6, print.output = 0, save.SNN = , force.recalc = T)
# scData <- StashIdent(object = scData, save.name = "CellType")
# scData <- RunTSNE(object = scData, dims.use = 1:75, do.fast = T, perplexity = 60)
# #
# scData <- SetAllIdent(object = scData, id = "CellType")

```

# Final preprocessing

Since there are several independent batches of samples with partially very different outcomes in terms of processing quality and cell viability we will use data set integration (Stuart and Butler et al., Cell 2019, doi: 10.1016/j.cell.2019.05.031) as implemented in Seurat v3.


```{r}
old_meta = scData@meta.data
scData = all_ge_data_filtered

scData@meta.data$batch = unlist(sapply(strsplit(scData@meta.data$sample_id_final, " / "), `[`,1))
scData@meta.data$Phase = old_meta$Phase
scData@meta.data$S.Score = old_meta$S.Score
scData@meta.data$G2M.Score = old_meta$G2M.Score

sc.list <- SplitObject(scData, split.by = "batch")

sc.list <- lapply(X = sc.list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
```



```{r}
sc.anchors = FindIntegrationAnchors(object.list = sc.list, dims = 1:20)
gc()
scData.combined <- IntegrateData(anchorset = sc.anchors, dims = 1:20)
gc()
```


```{r}
DefaultAssay(scData.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
#scData.combined <- ScaleData(scData.combined, verbose = TRUE, vars.to.regress = c("sample_id_final", "G2M_score", "G1_score"))
#scData.combined <- ScaleData(scData.combined, verbose = TRUE, vars.to.regress = c("percent_mito"))
scData.combined <- ScaleData(scData.combined, verbose = TRUE)
gc()
scData.combined <- RunPCA(scData.combined, npcs = 30, verbose = TRUE)
gc()
# t-SNE and Clustering
scData.combined <- RunUMAP(scData.combined, reduction = "pca", dims = 1:20)
scData.combined <- FindNeighbors(scData.combined, reduction = "pca", dims = 1:20)
scData.combined <- FindClusters(scData.combined, resolution = 0.5)
```

```{r}
save(scData.combined, file=file.path(data_storage_folder, "BCN_NAFLD.combined.RData"))
```

# Global clusters

```{r, fig.width=14}
DimPlot(scData.combined, reduction = "umap", group.by = "batch")

p1 <- DimPlot(scData.combined, reduction = "umap", group.by = "group")
p2 <- DimPlot(scData.combined, reduction = "umap", label = TRUE)
plot_grid(p1, p2)

DimPlot(scData.combined, reduction = "umap", split.by = "group", label = T)
```

## Cluster assignments

```{r}
m = table(scData.combined@meta.data$seurat_clusters, scData.combined@meta.data$sample_label)
kable(m)
```


```{r, fig.height=10, fig.width=8}
mm = as.matrix.data.frame(m)
colnames(mm) = colnames(m)
rownames(mm) = rownames(m)
mm_rel = sweep(mm, 2, apply(mm, 2,sum), "/")
pheatmap(mm_rel, main="Relative single cell contribution of each cluster to condition totals")
```

# Software versions

```{r}
sessionInfo()
```

