---
title: "Gene visualizations EGFL7 and possible interaction partners"
author: "Hilmar Berger"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---

```{r}
library(Seurat)
library(ggplot2)


result_folder = paste("./Results",format(Sys.time(), "%Y-%m-%d"),sep="/")
if (!file.exists(result_folder)) dir.create(result_folder, recursive=T)

#data_storage_folder = file.path("./Results/data_storage","2022-05-23")
#if (!file.exists(data_storage_folder)) dir.create(data_storage_folder, recursive=T)

```


# Selected genes

  - Notch1; Notch2 ; Notch3 & Notch4
  - EGFR
  - αvβ3 integrin (ITGAV + ITGB3)
  - also interacts with DLL4, Jagged1 (PMID: 31672772,  22160377)

```{r}
sel_genes = c("EGFL7","EGFR","ERBB2","ERBB3","ERBB4","ITGAV","ITGB3","NOTCH1","NOTCH2","NOTCH3","NOTCH4", "DLL4","JAG1")
sel_genes
```

In all dotplots, gene expression levels are centered to mean zero and unit variance for each gene to make them comparable

# BCN alone, all cells

```{r}
load(file.path("./Results/data_storage/2021-05-31", "BCN_NAFLD.combined_cluster_annotated.RData"))
```

```{r}
new_groups_text = "SampleID\tSAMPLE ID_BCN\tNAME\tstage\tstatus
AP8596\tpatient1\tpatient 1\tNAFLD-ADVANCED FIBROSIS\tNAFLD
AP9962\tAP003\tpatient 2\tNAFLD-ADVANCED FIBROSIS\tNAFLD
AR9955\tP181-control1\tpatient 4\tNAFLD-F0-F1\tNAFLD
AS2023\tP197\tpatient 3\tNAFLD-F0-F2\tNAFLD
AR9367\tP184\tcontrol 1\tOBESE\tOBESE
AR9368\tP189\tcontrol 2\tOBESE\tOBESE"
new_groups = read.table(text=new_groups_text, sep="\t", header=T)
new_groups$stage = gsub("-","_", new_groups$stage)
new_groups
rownames(new_groups) = new_groups$SampleID

scData.combined$stage_new = new_groups[scData.combined$orig.ident, "stage"]
scData.combined$status_new = new_groups[scData.combined$orig.ident, "status"]
```

Integrin alphaVbeta3 = 	ITGAV + ITGB3

```{r, fig.width=14, fig.height=10}

DotPlot(scData.combined, features = sel_genes, cols = "RdYlGn", assay = "RNA")
DotPlot(scData.combined, features = sel_genes, cols = "RdYlGn", assay = "RNA", split.by="status_new") + ggtitle("Split by patient group (obese controls vs. NAFLD)")
```

# BCN alone, only myeloid cells

```{r}
load(file.path("./Results/data_storage/2021-05-31", "BCN_NAFLD_MoMF_DC_only.combined_cluster_annotated.RData"))
```

```{r}
scData.combined$stage_new = new_groups[scData.combined$orig.ident, "stage"]
scData.combined$status_new = new_groups[scData.combined$orig.ident, "status"]
```


Integrin alphaVbeta3 = 	ITGAV + ITGB3

```{r, fig.width=14, fig.height=10}
DotPlot(scData.combined, features = sel_genes, cols = "RdYlGn", assay = "RNA")
DotPlot(scData.combined, features = sel_genes, cols = "RdYlGn", assay = "RNA", split.by="status_new") + ggtitle("Split by patient group (obese controls vs. NAFLD)")
```

# BCN + Hildgreth, all cells

```{r}
load(file.path("./Results/data_storage/2022-05-23", "BCN_NAFLD_Hildreth_complete.combined_cluster_annotated.RData"))
```



```{r, fig.width=14, fig.height=10}
DotPlot(scData.combined, features = sel_genes, cols = "RdYlGn", assay = "RNA")
DotPlot(scData.combined, features = sel_genes, cols = "RdYlGn", assay = "RNA", split.by = "DS") + ggtitle("Split by data set")
```

```{r, fig.width=10,fig.height=5}

for( g in sel_genes) {
  p = VlnPlot(scData.combined, assay = "RNA", features = g) + NoLegend()
  print(p)
}

```


# BCN + Hildgreth, only myeloid cells

```{r}
load(file.path("./Results/data_storage/2022-05-23", "scData_NAFLD_Hildreth_complete_MoMF.DC.cluster_annotated.RData"))
```


```{r, fig.width=10, fig.height=6}
DotPlot(scData_MoMF.DC, features = sel_genes, cols = "RdYlGn", assay = "RNA", scale=F)
```


# BCN + Ramachandran Liver, all cells

```{r}
load(file.path("./Results/data_storage/2022-01-31", "BCN_NAFLD_Ramachandran_complete.combined_cluster_annotated.RData"))
```


```{r, fig.width=14, fig.height=10}
DotPlot(scData.combined, features = sel_genes, cols = "RdYlGn", assay = "RNA")
DotPlot(scData.combined, features = sel_genes, cols = "RdYlGn", assay = "RNA", split.by = "DS") + ggtitle("Split by data set")
```

```{r, fig.width=14, fig.height=5}
for( g in sel_genes) {
  p = VlnPlot(scData.combined, assay = "RNA", features = g, split.by = "DS")# + NoLegend()
  print(p)
}
```



